---
title: "Method Comparison Graphs"
author: "Gabriel J. Odom, PhD, ThD"
date: "July 9, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Setup

## Load Data and Packages
```{r load, message = FALSE, warning = FALSE}
library(DMRcompare)
library(tidyverse)
library(ggpubr)
data("dmrcateRes_df")
data("probeLassoRes_df")
data("bumphunterRes_df")
data("combpRes_df")
```


## Subset to the Best-Performing Cases
```{r bestCases}
results_ls <- list()
results_ls$DMRcate <- 
  dmrcateRes_df %>%  
  filter(lambda == 500) %>% 
  filter(C == 5) %>% 
  mutate(method = paste0(method, "_lambda500_C5")) %>% 
  select(-one_of("lambda", "C"))

results_ls$ProbeLasso <- 
  probeLassoRes_df %>% 
  filter(adjPval == 0.05) %>% 
  filter(mLassoRad == 1000) %>% 
  filter(minDmrSep == 1000) %>% 
  mutate(method = paste0(method, "_adjP0.05_mLassR1000")) %>% 
  select(-one_of("adjPval", "mLassoRad", "minDmrSep"))

results_ls$Bumphunter <- 
  bumphunterRes_df %>% 
  filter(cutoffQ == 0.95) %>% 
  filter(maxGap == 250) %>% 
  mutate(nCPG_q1 = as.character(nCPG_q1)) %>%
  mutate(nCPG_med = as.character(nCPG_med)) %>%
  mutate(nCPG_q3 = as.character(nCPG_q3)) %>% 
  mutate(method = paste0(method, "_cutQ0.95_maxGap250")) %>% 
  select(-one_of("cutoffQ", "maxGap"))

results_ls$Comb_p <- 
  combpRes_df %>% 
  filter(combSeed == 0.05) %>% 
  filter(combDist == 750) %>%
  mutate(method = paste0(method, "_Seed0.05_Dist750")) %>% 
  select(-one_of("combSeed", "combDist"))
```


## Clean and Join
```{r cleanJoin}
results_df <- 
  results_ls %>% 
  bind_rows() %>% 
  filter(delta > 0) %>% 
  mutate(nCPG_q1 = as.numeric(nCPG_q1)) %>%
  mutate(nCPG_med = as.numeric(nCPG_med)) %>%
  mutate(nCPG_q3 = as.numeric(nCPG_q3)) %>% 
  mutate(method = factor(method)) %>% 
  select(method, delta, seed, TP, FP, FN, power, precision,
         AuPR, mcc, F1, nCPG_q1, nCPG_med, nCPG_q3)

resultsAve_df <- 
  results_df %>% 
  group_by(method, delta) %>%
  summarise(
    TP       = mean(TP),
    FP       = mean(FP),
    FN       = mean(FN),
    Pwr      = mean(power),
    Precis   = mean(precision),
    AuPR     = mean(AuPR),
    MCC      = mean(mcc),
    F1       = mean(F1),
    nCPG_q1  = mean(nCPG_q1),
    nCPG_med = mean(nCPG_med),
    nCPG_q3  = mean(nCPG_q3)
  )

results_df %>% 
  group_by(method) %>%
  summarise(med_nCPGs = median(nCPG_med, na.rm = TRUE)) %>% 
  knitr::kable()
```


# Figure 3: Power and Precision Plots by Delta

## Power
```{r powerXdelta}
power_gg <- 
  ggplot(data = resultsAve_df) + 
  theme_bw() +
  theme(panel.grid.minor = element_blank()) + 
  aes(x = delta, y = Pwr, group = method, colour = method) + 
  scale_y_continuous("Power", limits = c(0, 1)) +
  scale_x_continuous("Treatment Effect",
                     breaks = unique(results_df$delta),
                     labels = prettyNum(unique(results_df$delta)),
                     limits = c(0, 0.4)) +
  ggtitle("Power under Best-Performing Parameter Set vs. Treatment Effect") + 
  scale_color_discrete(name = "DMR Method",
                       breaks = levels(results_df$method),
                       labels = c("Bumphunter", "Comb-p",
                                  "DMRcate", "ProbeLasso")) +
  geom_line(size = 1)

power_gg
```


## Precision
```{r precisionXdelta, warning = FALSE}
precis_gg <- 
  ggplot(data = resultsAve_df) + 
  theme_bw() +
  theme(panel.grid.minor = element_blank()) + 
  aes(x = delta, y = Precis, group = method, colour = method) + 
  scale_y_continuous("Precision", limits = c(0.0, 1)) +
  scale_x_continuous("Treatment Effect",
                     breaks = unique(results_df$delta),
                     labels = prettyNum(unique(results_df$delta)),
                     limits = c(0, 0.4)) +
  ggtitle("Precision under Best-Performing Parameter Set vs. Treatment Effect") + 
  scale_color_discrete(name = "DMR Method",
                       breaks = levels(results_df$method),
                       labels = c("Bumphunter", "Comb-p",
                                  "DMRcate", "ProbeLasso")) +
  geom_line(size = 1)
precis_gg
```

## Factor of Delta
```{r deltaFct}
nCPGsFct_df <-
  resultsAve_df %>% 
  mutate(delta = factor(delta,
                        levels = unique(delta),
                        ordered = TRUE))
```

## Figure 4: Side-by-Side Precision and Power
```{r sideBYside}
power2_gg <- 
  ggplot(data = nCPGsFct_df) + 
  theme_bw() +
  theme(panel.grid.minor = element_blank()) + 
  aes(x = delta, y = Pwr, group = method, colour = method) + 
  scale_y_continuous("Power", limits = c(0, 1)) +
  scale_x_discrete(expression(mu),
                   breaks = unique(results_df$delta),
                   labels = prettyNum(unique(results_df$delta))) +
  scale_color_discrete(name = "DMR Method",
                       breaks = levels(results_df$method),
                       labels = c("Bumphunter", "Comb-p",
                                  "DMRcate", "ProbeLasso")) +
  ggtitle(" Power under Best-Performing \n Parameter Set vs. Treatment Effect") + 
  geom_line(size = 1)

precis2_gg <- 
  ggplot(data = nCPGsFct_df) + 
  theme_bw() +
  theme(panel.grid.minor = element_blank()) + 
  aes(x = delta, y = Precis, group = method, colour = method) + 
  scale_y_continuous("Precision", limits = c(0, 1)) +
  scale_x_discrete(expression(mu),
                   breaks = unique(results_df$delta),
                   labels = prettyNum(unique(results_df$delta))) +
  ggtitle(" Precision under Best-Performing \n Parameter Set vs. Treatment Effect") + 
  scale_color_discrete(name = "DMR Method",
                       breaks = levels(results_df$method),
                       labels = c("Bumphunter", "Comb-p",
                                  "DMRcate", "ProbeLasso")) +
  geom_line(size = 1)

ggarrange(precis2_gg, power2_gg,
          nrow = 1, ncol = 2,
          common.legend = TRUE,
          legend = "bottom")
```


# Box-and-Whiskers Plot of nCPGs
```{r med_and_IQR_nCPGs, warning = FALSE}
nCPGs_gg <-
  ggplot(data = nCPGsFct_df) + 
  theme_bw() +
  aes(x = delta, group = method, colour = method) +
  ggtitle("Median and IQR of DMR Sizes for Each Method under Optimal Parameter Settings") +
  scale_color_discrete(name = "DMR Method",
                       breaks = levels(nCPGsFct_df$method),
                       labels = c("Bumphunter", "Comb-p",
                                  "DMRcate", "ProbeLasso")) +
  scale_x_discrete("Treatment Effect") + 
  scale_y_continuous("Size of DMRs Detected") +
  geom_errorbar(
    mapping = aes(ymin = nCPG_q1, ymax = nCPG_q3),
    position = position_dodge(width = 0.5),
    width = 0.5,
    size = 1
  ) +
  geom_point(
    aes(y = nCPG_med),
    position = position_dodge(width = 0.5),
    size = 4
  )

nCPGs_gg
```


# Best-Case DMR Sizes

## Read and Compile Data
We will first read in the raw performance data for all four methods under their best-performing parameter setting. Zhen did not filter the original Comb-p results to `nCPGs`$ >= 5$, so we have to do that here.
```{r read_data_raw, warning = FALSE}
fileNames_char <- list.files("../extdata/best_cases_results")


nCPGs_ls <- lapply(fileNames_char, function(name){
  
  # Metadata
  name_char <- strsplit(name, split = "_")[[1]]
  
  # data
  data_ls <- readRDS(paste0("../extdata/best_cases_results/", name))
  data_df <- data_ls[[1]]
  if(name_char[1] == "CombpResults"){
    
    data_df <- data_df[, c("chrom", "start", "end",
                           "z_p", "n_probes")]
    colnames(data_df) <- c("dmr.chr", "dmr.start", "dmr.end",
                           "dmr.pval", "dmr.n.cpgs")
    data_df <- data_df[data_df$dmr.n.cpgs > 4, ]
    
  } else {
    data_df <- data_df[, c("dmr.chr", "dmr.start", "dmr.end",
                           "dmr.pval", "dmr.n.cpgs")]
  }
  
  # bind
  data_df$method <- gsub(pattern = "Results", replacement = "", name_char[1])
  data_df$delta  <- as.numeric(
    gsub(pattern = "delta", replacement = "", name_char[2])
  ) 
  data_df$seed   <- as.numeric(
    gsub(pattern = "seed", replacement = "", name_char[3])
  )
  
  # return
  data_df
  
})

nCPGs_df <- 
  nCPGs_ls %>% 
  bind_rows() %>% 
  select(delta, method, seed, everything()) %>% 
  arrange(delta) %>% 
  mutate(delta = factor(delta,
                        levels = unique(delta),
                        ordered = TRUE))
nCPGs_df %>% 
  group_by(method) %>%
  summarise(med_nCPGs = median(dmr.n.cpgs, na.rm = TRUE)) %>% 
  knitr::kable()
```

## Figure 6: Build Boxplot
Now we have a data frame with the number of CPGs for each method by each design point. We use `theme_classic()` to pretty-up the hidden outliers; there are quite a instances where the ProbeLasso method returned large DMR sizes.
```{r nCPGs_boxplot}
ggplot(data = nCPGs_df) +
  theme_classic() +
  theme(legend.position = "bottom") +
  aes(x = delta, y = dmr.n.cpgs, fill = method) +
  ggtitle("DMR Sizes under Best-Performing Parameter Settings for Considered Methods") +
  scale_y_continuous("Number of CpGs in Significant DMRs",
                     limits = c(5, 25)) +
  scale_x_discrete(expression(mu)) +
  scale_fill_discrete("DMR Method") +
  geom_boxplot(position = position_dodge(0.7),
               width = 0.5,
               outlier.color = "white")
```

## Violin Plot
This shows the very long tail for ProbeLasso
```{r nCPGs_violin}
ggplot(data = nCPGs_df) +
  theme_bw() + 
  aes(x = delta, y = dmr.n.cpgs, fill = method) +
  scale_y_continuous("Size of DMRs Detected") +
  scale_x_discrete(expression(mu)) +
  scale_fill_discrete("DMR Method") +
  geom_violin()
```
